\huge{Comment on `The Limits of Cosmic Shear'}

### 2016-12-06 Limber_Comment.tex
included comments from KK

In this comment paper we discuss the flat sky and Limber approximations that can be used when deriving theoretical predictions for the principally used cosmic shear statistic, $\xi_\pm$, the two-point shear correlation function. The Limber approximation for cosmic shear has been previously explored by \citet{giannantonio/etal:2012} finding that it provided sufficient accuracy for future cosmic shear studies. Recently, however, \citet{kitching/etal:2016} reported that when considering the flat sky and Limber approximations in concert, tensions between weak lensing and CMB experiments can be reduced. This was demonstrated with a re-analysis of the non-tomographic weak lensing data from the CFHTLenS survey \citep{kilbinger/etal:2013}.

@@

In section~\ref{sec:photoz} we review our current knowledge of photometric redshift uncertainties and the sophisticated mitigation strategies that have already been applied to data.  We suggest that a possible misinterpretation in the direction of the redshift biases reported in \citet{choi/etal:2016} has lead to the incorrect conclusion of \citet{kitching/etal:2016} that these redshift biases can resolve the tension with Planck.

>>>>

In this comment paper we discuss the flat sky and Limber approximations that can be used when deriving theoretical predictions for the principally used cosmic shear statistic, $\xi_\pm$, the two-point shear correlation function. The Limber approximation for cosmic shear has been previously explored by \citet{giannantonio/etal:2012} finding that it provided sufficient accuracy for future cosmic shear studies. Recently, however, \citet{kitching/etal:2016} reported that when considering the flat sky and Limber approximations in concert, tensions between weak lensing and CMB experiments can be reduced. They demonstrate the effect with a re-analysis of the non-tomographic weak lensing data from the CFHTLenS survey \citep{kilbinger/etal:2013}.

@@

In section~\ref{sec:photoz} we review our current knowledge of photometric redshift uncertainties and the sophisticated mitigation strategies that have already been applied to data.  We suggest that a possible misinterpretation in the sign of the redshift biases reported in \citet{choi/etal:2016} has led to the incorrect conclusion of \citet{kitching/etal:2016} that these redshift biases can resolve the tension with Planck.

### 2016-12-06 cfhtlens.tex
included comments from KK

The Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS) represented a major step forward for the field of weak gravitational lensing, in terms of improved accuracy in data reduction \citep{erben/etal:2013}, the implementation of Gaussianised matched multi-band photometry \citep{hildebrandt/etal:2012}, cross-correlation clustering analysis between photometric redshift slices to verify tomographic redshift distributions \citep{benjamin/etal:2013}, accurate calibrated shape measurements \citep{miller/etal:2013} and a full suite of informative systematic tests to select a clean data set \citep{heymans/etal:2012}.    Since the public release of this survey in 2013, the community has continued to scrutinise and advance our understanding of CFHTLenS by identifying a number of areas where analyses could improve:
\begin{itemize}
\item{\citet{choi/etal:2016} identified biases in the tomographic photometric redshift distributions using a more effective clustering analysis, in comparison to \citet{benjamin/etal:2013}, by incorporating newly overlapping spectroscopy from the Sloan Digital Sky Survey.  The CFHTLenS tomographic cosmological analysis was then revisited by \citet{joudaki/etal:2016} in order to include a full redshift error analysis based on the results from \citet{choi/etal:2016}, discussed further in section~\ref{sec:photoz}.}
\item{\citet{asgari/etal:2016} used the stringent `COSEBI' statistic to identify significant non-lensing `B-mode' distortions when the CFHTLenS data was split into tomographic slices.}
\item{\citet{kuijken/etal:2015} showed that the CFHTLenS shear calibration corrections derived in \citet{miller/etal:2013} were underestimated as a result of an imperfect match between the galaxy populations in the data and image simulations.}
\item{\citet{fenechconti/etal:2016} demonstrated that the CFHTLenS data would have been subject to a weight bias that favours galaxies that are more intrinsically oriented with the point-spread function.  They also showed that the impact of calibration selection biases, that were not considered in \citet{miller/etal:2013}, would have lead to the over-correction of multiplicative shear bias in the CFHTLenS analyses, by a few percent.}
\item{\citet{joudaki/etal:2016} updated the CFHTLenS covariance matrices using larger-box numerical simulations that were less subject to the lack of power on large scales.}
\end{itemize}
All these advances in our understanding were incorporated and accounted for in the recent KiDS cosmic shear analysis \citep{hildebrandt/etal:2016} which reports a $2.3 \sigma$ tension with Planck.  Efforts are now underway to fully re-analyse CFHTLenS using the advanced KiDS analysis pipeline with revised shape measurements and calibrations for the shear and photometric redshifts.  Given the known shortcomings with the original CFHTLenS results, which will impact in different ways on the cosmological conclusions that one can draw from the survey, it is surprising that \citet{kitching/etal:2016} chose to focus their study on the cosmic shear analysis of \citet{kilbinger/etal:2013}.

\citet{kitching/etal:2016} choose to only vary $\sigma_8$ in their analysis, fixing all other parameters to their Planck values. They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) of the data from \citet{kilbinger/etal:2013}.    We argue that this reduction in the error by a factor of $\sim 2$, in comparison to a full six-parameter fit, could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.  Nevertheless, we repeated the `one-parameter' analysis for the full range of Limber cases listed in Table~\ref{tab:Tl_nu} finding the lowest conditional value of $\sigma_8 = 0.743 \pm 0.023$ for the `Standard' (KSt) case and the highest conditional value of $\sigma_8 = 0.744 \pm 0.023$ for the `Baseline Limber Flat` (LF) case.  As expected from the comparison in Figure~\ref{fig:Cl_xi} the different combinations of assuming flat or spherical sky, baseline or extended Limber, make little difference to the cosmological parameter constraints, even in this `one-parameter' analysis.

@@

\subsection{On the application to CMB lensing}
\citet{kitching/etal:2016} argue that their correction to the Limber and flat sky approximations would also affect predictions involving the shear and convergence cross-spectrum, such that CMB lensing - galaxy lensing cross-correlation should include a pre-term $ [\ell(\ell+1)]^2/(\ell+ 0.5)^4.$.   We have verified that this term would have negligible impact on the current measurements, given the relatively low signal to noise ratio. For example, most measurements \citep{hand/etal:2015, liu/hill:2015, kirk/etal:2016,harnois-deraps/etal:2016} are carried over angular multipoles in the range $\ell \in [20-2000]$, and are organized in $\sim$ five bins. The size of the pre-term $ [(\ell(\ell+1))^2]/(\ell+ 0.5)^4.$, averaged over the first bin, is of the order of 2\%, and then becomes sub-percent for all other bins. In comparison, the size of the statistical error bars are of the order 25-35\% with additional uncertainty due to photometric redshifts (10-15\%) and intrinsic galaxy alignments (10-15\%).

>>>>

The Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS) represented a major step forward for the field of weak gravitational lensing, in terms of improved accuracy in data reduction \citep{erben/etal:2013}, the implementation of PSF-Gaussianised matched multi-band photometry \citep{hildebrandt/etal:2012}, cross-correlation clustering analysis between photometric redshift slices to verify tomographic redshift distributions \citep{benjamin/etal:2013}, accurate calibrated shape measurements \citep{miller/etal:2013} and a full suite of informative systematic tests to select a clean data set \citep{heymans/etal:2012}.    Since the public release of this survey in 2013, the community has continued to scrutinise and advance our understanding of CFHTLenS by identifying a number of areas where analyses could improve:
\begin{itemize}
\item{\citet{choi/etal:2016} identified biases in the tomographic photometric redshift distributions using a more effective clustering analysis, in comparison to \citet{benjamin/etal:2013}, by incorporating newly overlapping spectroscopy from the Sloan Digital Sky Survey.  The CFHTLenS tomographic cosmological analysis was then revisited by \citet{joudaki/etal:2016} in order to include a full redshift error analysis based on the results from \citet{choi/etal:2016}, discussed further in section~\ref{sec:photoz}.}
\item{\citet{asgari/etal:2016} used the stringent `COSEBI' statistic to identify significant non-lensing `B-mode' distortions when the CFHTLenS data was split into tomographic slices.}
\item{\citet{kuijken/etal:2015} showed that the CFHTLenS shear calibration corrections derived in \citet{miller/etal:2013} were underestimated as a result of an imperfect match between the galaxy populations in the data and image simulations.}
\item{\citet{fenechconti/etal:2016} demonstrated that the CFHTLenS data would have been subject to a weight bias that favours galaxies that are more intrinsically oriented with the point-spread function.  They also showed that the impact of calibration selection biases, that were not considered in \citet{miller/etal:2013}, would have lead to the over-correction of multiplicative shear bias in the CFHTLenS analyses, by a few percent.}
\item{\citet{joudaki/etal:2016} updated the CFHTLenS covariance matrices using larger-box numerical simulations that were less subject to the lack of power on large scales.}
\end{itemize}
All these advances in our understanding were incorporated and accounted for in the recent KiDS cosmic shear analysis \citep{hildebrandt/etal:2016} which reports a $2.3 \sigma$ tension with Planck.  Efforts are now underway to fully re-analyse CFHTLenS using the advanced KiDS analysis pipeline with revised shape measurements and calibrations for the shear and photometric redshifts.  Given the known shortcomings with the original CFHTLenS results, which will impact in different ways on the cosmological conclusions that one can draw from the survey, it is surprising that \citet{kitching/etal:2016} chose to focus their study on the cosmic shear analysis of \citet{kilbinger/etal:2013}.

\citet{kitching/etal:2016} choose to only vary $\sigma_8$ in their analysis, fixing all other parameters to their Planck values. They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) of the data from \citet{kilbinger/etal:2013}.    We argue that this reduction in the error by a factor of $\sim 2$, in comparison to a full six-parameter fit, could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.  Nevertheless, we repeated the `one-parameter' analysis for the full range of Limber cases listed in Table~\ref{tab:Tl_nu} and are unable to confirm their results.  We find the lowest conditional value of $\sigma_8 = 0.743 \pm 0.023$ for the `Standard' (KSt) case and the highest conditional value of $\sigma_8 = 0.744 \pm 0.023$ for the `Baseline Limber Flat` (LF) case.  As expected from the comparison in Figure~\ref{fig:Cl_xi} the different combinations of assuming flat or spherical sky, baseline or extended Limber, make little difference to the cosmological parameter constraints, even in this `one-parameter' analysis.

@@

\subsection{On the application to CMB lensing}
\citet{kitching/etal:2016} argue that their correction to the Limber and flat sky approximations would also affect predictions involving the shear and convergence cross-spectrum, such that CMB lensing - galaxy lensing cross-correlation should include a pre-term $ [\ell(\ell+1)]^2/(\ell+ 0.5)^4$.   We have verified that this term would have negligible impact on current measurements, given the relatively low signal to noise ratio. For example, most measurements \citep{hand/etal:2015, liu/hill:2015, kirk/etal:2016,harnois-deraps/etal:2016} are carried over angular multipoles in the range $\ell \in [20-2000]$, and are organized in $\sim$ five bins. The size of the pre-term $ [(\ell(\ell+1))^2]/(\ell+ 0.5)^4$, averaged over the first bin, is of the order of 2\%, and then becomes sub-percent for all other bins. In comparison, the size of the statistical error bars are of the order 25-35\% with additional uncertainty due to photometric redshifts (10-15\%) and intrinsic galaxy alignments (10-15\%).

### 2016-12-06 conclusion.tex
included comments from KK

The treatment of photo-$z$ uncertainties in \cite{kitching/etal:2016} is based on a simple linear shift of redshift distributions. This has been shown to be insufficient to capture the full complexity of the problem. More importantly though, it seems like \cite{kitching/etal:2016} apply the photo-$z$ calibrations from \cite{choi/etal:2016} in the wrong direction. Instead of alleviating the tension between CFHTLenS and Planck a correct application of the offsets would increase the discrepancy, in agreement with a recent re-analysis of the CFHTLenS data \citep{joudaki/etal:2016} as well as results from the KiDS project \citep{hildebrandt/etal:2016}.

>>>>

The treatment of photo-$z$ uncertainties in \cite{kitching/etal:2016} is based on a simple linear shift of redshift distributions. This has been shown to be insufficient to capture the full complexity of the problem.  More importantly though, we suggest that \cite{kitching/etal:2016} have applied the photo-$z$ calibrations from \cite{choi/etal:2016} in the wrong sense. Instead of alleviating the tension between CFHTLenS and Planck a correct application of the offsets would increase the discrepancy, as shown with a recent re-analysis of the CFHTLenS data \citep{joudaki/etal:2016} as well as results from the KiDS project \citep{hildebrandt/etal:2016}.

### 2016-12-06 photoz.tex
included comments from KK

\citet{choi/etal:2016} additionally showed that applying a simple linear shift to the photometric redshift distributions was insufficient when modelling the true underlying redshift distribution.  In \citet{hildebrandt/etal:2016}, a direct calibration method was used to determine the redshift distribution of four tomographic bins.  Multiple bootstrap samples of the resulting calibrated distribution allowed for the characterisation of both the uncertainty in the mean and shape of the redshift distribution.  The inclusion of this more sophisticated treatment of photometric redshift errors did not alleviate the tension between Planck and KiDS parameter constraints.

>>>>

\citet{choi/etal:2016} additionally showed that applying a simple linear shift to the photometric redshift distributions was insufficient when modelling the true underlying redshift distribution.  In \citet{hildebrandt/etal:2016}, a direct calibration method was used to determine the redshift distribution of four tomographic bins.  Multiple bootstrap samples of the resulting calibrated distribution allowed for the characterisation of both the uncertainty in the mean and shape of the redshift distribution.  The inclusion of this more sophisticated treatment of photometric redshift errors was not found to alleviate the tension between Planck and KiDS parameter constraints.

### 2016-12-06 theory.tex
included comments from KK

 \subsection{Cosmic shear without the Limber Approximation}
In collecting our comments on \citet{kitching/etal:2016} we contacted colleagues who had previously tested the impact of the Limber approximation on cosmic shear studies by comparing a Limber approximated cosmic shear power spectrum with an exact calculation.  The majority of these analyses, carried out many years ago, remained unpublished as no significant deviation was found.  One notable exception, however, is \citet{giannantonio/etal:2012} who find that for a cosmic shear power spectrum, the extended Limber approximation is consistent with the exact calculation for $\ell>5$, with any differences well within cosmic-variance errors for a Euclid-like survey\footnote{\citet{kitching/etal:2016} comment on this result, but indicate that it is in error owing to the application of a fixed low-$\ell$ limit.  \citet{giannantonio/etal:2012} however state that this limit is only applied in a positive curvature case.}.  We therefore argue that the assertion by \citet{kitching/etal:2016} that the majority of cosmic shear data analyses to date have made an `axiomatic assumption (i.e. an unquestioned and untested assumption at the beginning of the analysis)'  is unwarranted.  Not only is there published work that has questioned and tested the Limber approximation for cosmic shear studies, individual teams have also verified this result internally.

Given the insignificant effect of using the Limber approximation found by various groups, the ability to calculate an exact non-Limber solution has unfortunately not been maintained in theoretical code over the years.  Instead these codes have evolved to account for important approximations such as the impact of baryon feedback on the matter power spectrum and non-zero neutrino masses \citep[see for example][]{joudaki/etal:2016, mead/etal:2016}.    We are therefore unable to readily determine non-Limber theoretical calculations for the shear-correlation function to directly compare with the results of \citet{kitching/etal:2016}.   Instead we refer the reader to Figure 3 of \citet{giannantonio/etal:2012} to draw their own conclusions.

>>>>

 \subsection{Cosmic shear without the Limber Approximation}
In collecting our comments on \citet{kitching/etal:2016} we contacted colleagues who had previously tested the impact of the Limber approximation on cosmic shear studies by comparing a Limber approximated cosmic shear power spectrum with an exact calculation.  The majority of these analyses, carried out many years ago, remained unpublished in peer reviewed journals as no significant deviation was found\footnote{See for example appendix M of the 2010 PhD thesis from Donghui Jeong (\url{http://www.personal.psu.edu/duj13/dissertation/djeong_diss_appM.pdf}) and Chapter 7 of 2012 PhD thesis from Nicolas Van de Rijt (\url{http://ipht.cea.fr/Docspht/articles/t12/080/public/these_vanderijt2012.pdf}).}.  One notable exception, however, is \citet{giannantonio/etal:2012} who find that for a cosmic shear power spectrum, the extended Limber approximation is consistent with the exact calculation for $\ell>5$, with any differences well within cosmic-variance errors for a Euclid-like survey\footnote{\citet{kitching/etal:2016} comment on this result, but indicate that it is in error owing to the application of a fixed low-$\ell$ limit.  \citet{giannantonio/etal:2012} however state that this limit is only applied in a positive curvature case.}.  We therefore argue that the assertion by \citet{kitching/etal:2016} that the majority of cosmic shear data analyses to date have made an `axiomatic assumption (i.e. an unquestioned and untested assumption at the beginning of the analysis)'  is unwarranted.  Not only is there published work that has questioned and tested the Limber approximation for cosmic shear studies, individual teams have also verified this result internally.

Given the insignificant effect of using the Limber approximation found by various groups, the ability to calculate an exact non-Limber solution has unfortunately not been maintained in theoretical code over the years.  Instead these codes have evolved to account for (much more) important approximations such as the impact of baryon feedback on the matter power spectrum and non-zero neutrino masses \citep[see for example][]{joudaki/etal:2016, mead/etal:2016}.    We are therefore unable to readily determine non-Limber theoretical calculations for the shear-correlation function to directly compare with the results of \citet{kitching/etal:2016}.   Instead we refer the reader to Figure 3 of \citet{giannantonio/etal:2012} to draw their own conclusions.
